Five-dimensional seismic data reconstruction using the optimally damped rank-reduction method

### 2020-11-03 odrr5d/paper.tex
2020年11月 2日 星期一 21时31分03秒 CST

We also compare the level-four block Hankel matrices of different methods. Figures \ref{fig:H} and \ref{fig:H0} shows the Hankel matrices for the clean and incomplete data for the frequency slice of 30Hz. Figures \ref{fig:H_z} and \ref{fig:H0_z} show the zoomed Hankel matrices for the clean and incomplete data. The zooming areas are highlighted by the red rectangles shown in Figures \ref{fig:H} and \ref{fig:H0}. Because of the missing traces, the Hankel matrix contains a number of blank areas. Figure \ref{fig:Hs} shows the Hankel matrices for different methods. The left column correspond to the Hankel matrices after 1 iteration. The right column correspond to the Hankel matrices after 10 iterations. We only conclude from Figure \ref{fig:Hs} that after 10 iterations all three methods obtain similar Hankel matrices as the exact solution shown in Figure \ref{fig:H} but we cannot see the difference between different methods clearly. The differences between different methods can be clearly observed from the zoomed Hankel matrices, as shown in Figure \ref{fig:Hsz}. It is clear that from the top to bottom in Figure \ref{fig:Hsz}, the Hankel matrix becomes smoother and smoother and is closer to the exact solution shown in Figure \ref{fig:H_z}.

In addition to the local similarity mentioned above, we also use the signal-to-noise ratio (SNR) defined as follows to quantitatively measure the performance \cite[]{weilin2016,yangkang2017sgk,weilin2017gji}:
\begin{equation}
\label{eq:dsnr}
SNR=10\log_{10}\frac{\Arrowvert \mathbf{s} \Arrowvert_2^2}{\Arrowvert \mathbf{s} -\hat{\mathbf{s}}\Arrowvert_2^2}.
\end{equation}
where $\mathbf{s}$ denotes the vectorized clean data and $\hat{\mathbf{s}}$ denotes the vectorized reconstructed data. In this example, the SNRs of the observed data, the reconstructed results via the RR method, the DRR method, and ORR method are -4.59 dB, 5.51 dB, 9.54 dB, and 11.83 dB, respectively. In this test, we use $rank=10$, assuming that we do not have a prior information about the structural complexity of the data and thus needing to set a relatively high rank. We then test the performance when the rank is chosen smaller. The reconstructed data when $rank=5$ are shown in Figure \ref{fig:syn_dn_N5}. Their corresponding error sections are shown in Figure \ref{fig:syn_e_N5}. From Figures \ref{fig:syn_dn_N5} and \ref{fig:syn_e_N5}, we can reach almost the same conclusion as the last test. In this case, the SNRs of the reconstructed results via the RR method, the DRR method, and ORR method are 8.01 dB, 10.87 dB, and 11.97 dB, respectively. The detailed SNR comparison is shown in Table \ref{tbl:plane}, including the case when $rank=3$.

@@

%
%\begin{figure}
%	\centering
%	\subfloat[]{\includegraphics[width=0.45\textwidth]{Fig/H}
%    \label{fig:H}}
%    \subfloat[]{\includegraphics[width=0.45\textwidth]{Fig/H0}
%    \label{fig:H0}}\\
%	\subfloat[]{\includegraphics[width=0.45\textwidth]{Fig/H_z}
%    \label{fig:H_z}}
%    \subfloat[]{\includegraphics[width=0.45\textwidth]{Fig/H0_z}
%    \label{fig:H0_z}}
%	\caption{Hankel matrices for the clean and observed data for frequency slice of 30Hz. (a) Hankel matrix of the clean data. (b) Hankel matrix of the observed data. (c) \& (d) Zoomed Hankel matrices of (a) \& (b). The zooming area is highlighted by the red rectangles.}
%	\label{fig:HH}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H1_1}
%    \label{fig:H1_1}}
%    \subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H2_1}
%    \label{fig:H2_1}}\\
%	\subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H1_2}
%    \label{fig:H1_2}}
%    \subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H2_2}
%    \label{fig:H2_2}}\\
%	\subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H1_3}
%    \label{fig:H1_3}}
%    \subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H2_3}
%    \label{fig:H2_3}}
%	\caption{Hankel matrices for different methods for frequency slice of 30Hz. Left column: Hankel matrix after 1 iteration. Right: Hankel matrix after 10 iterations. Top row: Hankel matrices for the RR method. Middle row: Hankel matrix for the DRR method. Bottom row: Hankel matrix for the ORR method.}
%	\label{fig:Hs}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H1_1_z}
%    \label{fig:H1_1_z}}
%    \subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H2_1_z}
%    \label{fig:H2_1_z}}\\
%	\subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H1_2_z}
%    \label{fig:H1_2_z}}
%    \subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H2_2_z}
%    \label{fig:H2_2_z}}\\
%	\subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H1_3_z}
%    \label{fig:H1_3_z}}
%    \subfloat[]{\includegraphics[width=0.4\textwidth]{Fig/H2_3_z}
%    \label{fig:H2_3_z}}
%	\caption{Zoomed Hankel matrices for different methods for frequency slice of 30Hz. Left column: Hankel matrix after 1 iteration. Right: Hankel matrix after 10 iterations. Top row: Hankel matrices for the RR method. Middle row: Hankel matrix for the DRR method. Bottom row: Hankel matrix for the ORR method.}
%	\label{fig:Hsz}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=0.8\textwidth]{Fig/syn_dn_N5}
%	\caption{Common offset gather comparison for the synthetic example (reshaped into a 2-D matrix). (a) Reconstructed data using the RR method. (b) Reconstructed data using the DRR method. (c) Reconstructed data using the ORR method.  In this case, $rank=5$.}
%	\label{fig:syn_dn_N5}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=0.8\textwidth]{Fig/syn_e_N5}
%	\caption{Common offset gather comparison for the synthetic example (reshaped into a 2-D matrix). (a) Reconstruction error using the RR method. (b) Reconstruction error using the DRR method. (c) Reconstruction error using the ORR method. In this case, $rank=5$. }
%	\label{fig:syn_e_N5}
%\end{figure}

>>>>

We also compare the level-four block Hankel matrices of different methods. Figures \ref{fig:H_clean} and \ref{fig:H_obs} shows the Hankel matrices for the clean and incomplete data for the frequency slice of 30Hz. Figures \ref{fig:H-z-clean} and \ref{fig:H-z-obs} show the zoomed Hankel matrices for the clean and incomplete data. The zooming areas are highlighted by the red rectangles shown in Figures \ref{fig:H_clean} and \ref{fig:H_obs}. Because of the missing traces, the Hankel matrix contains a number of blank areas. Figure \ref{fig:H_rr1,H_rr10,H_drr1,H_drr10,H_odrr1,H_odrr10} shows the Hankel matrices for different methods. The left column correspond to the Hankel matrices after 1 iteration. The right column correspond to the Hankel matrices after 10 iterations. We only conclude from Figure \ref{fig:H_rr1,H_rr10,H_drr1,H_drr10,H_odrr1,H_odrr10} that after 10 iterations all three methods obtain similar Hankel matrices as the exact solution shown in Figure \ref{fig:H_clean} but we cannot see the difference between different methods clearly. The differences between different methods can be clearly observed from the zoomed Hankel matrices, as shown in Figure \ref{fig:H-z-rr1,H-z-rr10,H-z-drr1,H-z-drr10,H-z-odrr1,H-z-odrr10}. It is clear that from the top to bottom in Figure \ref{fig:H-z-rr1,H-z-rr10,H-z-drr1,H-z-drr10,H-z-odrr1,H-z-odrr10}, the Hankel matrix becomes smoother and smoother and is closer to the exact solution shown in Figure \ref{fig:H-z-clean}.

In addition to the local similarity mentioned above, we also use the signal-to-noise ratio (SNR) defined as follows to quantitatively measure the performance \cite[]{weilin2016,yangkang2017sgk,weilin2017gji}:
\begin{equation}
\label{eq:dsnr}
SNR=10\log_{10}\frac{\Arrowvert \mathbf{s} \Arrowvert_2^2}{\Arrowvert \mathbf{s} -\hat{\mathbf{s}}\Arrowvert_2^2}.
\end{equation}
where $\mathbf{s}$ denotes the vectorized clean data and $\hat{\mathbf{s}}$ denotes the vectorized reconstructed data. In this example, the SNRs of the observed data, the reconstructed results via the RR method, the DRR method, and ORR method are -4.59 dB, 5.51 dB, 9.54 dB, and 11.83 dB, respectively. In this test, we use $rank=10$, assuming that we do not have a prior information about the structural complexity of the data and thus needing to set a relatively high rank. We then test the performance when the rank is chosen smaller. The reconstructed data when $rank=5$ are shown in Figure \ref{fig:figure10}. Their corresponding error sections are shown in Figure \ref{fig:figure11}. From Figures \ref{fig:figure10} and \ref{fig:figure11}, we can reach almost the same conclusion as the last test. In this case, the SNRs of the reconstructed results via the RR method, the DRR method, and ORR method are 8.01 dB, 10.87 dB, and 11.97 dB, respectively. The detailed SNR comparison is shown in Table \ref{tbl:plane}, including the case when $rank=3$.

@@

\inputdir{hankel}
\multiplot{4}{H_clean,H_obs,H-z-clean,H-z-obs}{width=0.45\textwidth}{Hankel matrices for the clean and observed data for frequency slice of 30Hz. (a) Hankel matrix of the clean data. (b) Hankel matrix of the observed data. (c) \& (d) Zoomed Hankel matrices of (a) \& (b). The zooming area is highlighted by the red rectangles.}

\multiplot{6}{H_rr1,H_rr10,H_drr1,H_drr10,H_odrr1,H_odrr10}{width=0.35\textwidth}{Hankel matrices for different methods for frequency slice of 30Hz. Left column: Hankel matrix after 1 iteration. Right: Hankel matrix after 10 iterations. Top row: Hankel matrices for the RR method. Middle row: Hankel matrix for the DRR method. Bottom row: Hankel matrix for the ORR method.}

\multiplot{6}{H-z-rr1,H-z-rr10,H-z-drr1,H-z-drr10,H-z-odrr1,H-z-odrr10}{width=0.35\textwidth}{Zoomed Hankel matrices for different methods for frequency slice of 30Hz. Left column: Hankel matrix after 1 iteration. Right: Hankel matrix after 10 iterations. Top row: Hankel matrices for the RR method. Middle row: Hankel matrix for the DRR method. Bottom row: Hankel matrix for the ORR method.}

\inputdir{synth2}
\plot{figure10}{width=0.8\textwidth}{Common offset gather comparison for the synthetic example (reshaped into a 2-D matrix). (a) Reconstructed data using the RR method. (b) Reconstructed data using the DRR method. (c) Reconstructed data using the ORR method.  In this case, $rank=5$. }
\plot{figure11}{width=0.8\textwidth}{Common offset gather comparison for the synthetic example (reshaped into a 2-D matrix). (a) Reconstruction error using the RR method. (b) Reconstruction error using the DRR method. (c) Reconstruction error using the ORR method. In this case, $rank=5$. }
